Learning Constraints and Optimization Criteria

### 2015-12-15 Paper.tex
Final content edits

\begin{itemize}
  \item \sam{Why interesting, connection to Campignotto, Domingos, Kok, Why preferences useful?}
  \item \sam{Real tasks with preferences? Small scale evaluation}
  \item \sam{Lot of background knowledge assumed (vs. constraint programming, probabilistic programming, demo model and need for learning)}
\end{itemize}

@@

We use weighted first-order logical theories to represent constrained optimization problems.
The weights capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
Furthermore, we learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
We use ILP to learn the clausal theories from interpretations, and preference learning to determine the weights.
Therefore, users provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraints and integrates constraint satisfaction and optimization.
It also extends the propositional approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
Preferences are useful in situation where the score of an example is hard to quantify and in interactive settings because it is often easier for humans to provide preferences over few examples at a time than absolute scores.

@@

Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
If a clause covers enough examples it is added to the result set, provided it is not entailed by other clauses already in the result set.
In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended (refined) by adding a new literal $l$.
The new clause, $c' = c \cup \{l\}$ is more general than $c$ ($c' \models c$) and potentially covers more examples.
By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
In order to do this efficiently, our implementation uses a pre-calculated list of atoms, given a maximal number of variables that can occur within a clause.
Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.

@@

\svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers a robust and efficient implementation.
The weights that it assigns to the features can be used directly as weights for the clauses.
While the current implementation uses \svm{}, other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.

@@

The learning process is parameterized by the maximal number of variables and literals allowed per clause.
If these limits are too large, constraints are found that over-fit the training data.
These constraints are too specific and will exclude valid solutions that are not in the training set.
However, if the chosen limits are too small, the necessary constraints will not be found.

Over-fitting can be addressed by removing constraints that exclude valid solutions (manually or automatically) as well as providing more (or larger) training examples.
Negative examples can be used to detect under-fitting, indicating that the parameter values are too small.

@@

Increasing the number of variables or literals per clause negatively impacts the efficiency.
Especially allowing both more variables \emph{and} more literals can steeply increase the execution time.
Including additional examples, however, only increases the execution time by a constant factor.
Since adding more examples can help improve the accuracy, this trade-off is often worthwhile.

@@

\section{Conclusion}
We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
Implementations for both clause discovery and clausal optimization have been provided that accomplish these tasks using first-order logic clauses.

>>>>

% \begin{itemize}
%   \item \sam{Why interesting, connection to Campignotto, Domingos, Kok, Why preferences useful?}
%   \item \sam{Real tasks with preferences? Small scale evaluation}
%   \item \sam{Lot of background knowledge assumed (vs. constraint programming, probabilistic programming, demo model and need for learning!)}
% \end{itemize}

@@

We use weighted first-order logical theories to represent constrained optimization problems (as illustrated in example~1).
The weights capture the optimization function, and (first-order) weighted MAX-SAT can be used to determine the \emph{best} solutions.
Furthermore, we learn such models by combining inductive logic programming (ILP) principles (such as clausal discovery \cite{DeRaedt:ClausalDiscovery}) with preference learning.
We use ILP to learn the clausal theories from interpretations, and preference learning to determine the weights.
Therefore, users provide examples of possible worlds and a set of preferences that specify which examples are preferred over other ones.
This setting extends traditional constraint learning and ILP approaches with the ability to learn soft constraints and integrates constraint satisfaction and optimization.
It also extends the propositional approach of \cite{campigotto2011active} as first-order theories are learned and preferences are used rather than examples with target values.
Preferences are useful in situations where the score of an example is hard to quantify (e.g. comparing songs or designs) and it is often easier for humans to provide preferences over few examples at a time than absolute scores.

@@

Starting with the empty clause ($\square$), clauses are refined until they cover enough examples.
If a clause covers enough examples it is added to the result set, unless it is entailed by clauses already in the result set.
In the refinement step, a clause $c = \{l_1, l_2, ..., l_k\}$ is extended (refined) by adding a new literal $l$.
The new clause, $c' = c \cup \{l\}$ is more general than $c$ ($c' \models c$) and potentially covers more examples.
By adding just one literal the refinement operator $\rho$ generates the set of clauses that are minimally more general.
In order to do this efficiently, our implementation uses a pre-calculated list of atoms, given a maximal number of variables that can occur within a clause.
Within the body and head of a clause, atoms may only be added in a specific order to avoid generating redundant clauses.

@@

\svm{} \cite{joachims2006training} was chosen to find the scoring function since it uses a linear model and offers a robust and efficient implementation.
The weights that it assigns to the features can be used directly as weights for the clauses.
Other linear ranking systems such as Coordinate Ascent \cite{metzler2007linear} could also be used.

@@

The learning process is parameterized by the maximal number of variables and literals allowed per clause.
If these limits are too large, constraints are found that over-fit the training data.
These constraints are too specific and will exclude valid solutions that are not in the training set.
Over-fitting can be addressed by removing invalid constraints (manually or automatically) and providing more (or larger) training examples.
On the other hand, if the chosen limits are too small, the necessary constraints will not be found.
Under-fitting can be detected using negative examples.

@@

Increasing the number of variables or literals per clause negatively impacts the efficiency.
Especially increasing both parameters can steeply increase the execution time.
The current implementation may still need significant time to solve large problems that include many predicates.
However, including additional or larger examples usually only increases the execution time by a constant factor.

@@

\section{Conclusion}
We presented an approach to automatically acquire constraints and optimization criteria from examples and preferences.
Implementations for both clause discovery and clausal optimization were provided that accomplish these tasks using first-order logic clauses.
